Arrows for Parallel Computations

### 2017-11-25 paper/content/benchmarks.tex
minor
best and worst benchmarks redone in one table for easier lookup

\subsection{Hardware}

In the following we describe the benchmarks of our parallel DSL and algorithmic skeletons we conducted. They were run both in a shared and in a distributed
memory setting. All benchmarks were done on the \enquote{Glasgow grid}, consisting of
16 machines with 2 Intel\SymbReg~Xeon\SymbReg~E5-2640 v2 and 64 GB of DDR3 RAM each. Each processor has 8 cores and 16 (hyper-threaded) threads with a base frequency of 2 GHz and a turbo frequency of 2.50 GHz. This results in a total of 256 cores and 512 threads for the whole cluster. The operating system was Ubuntu 14.04 LTS with Kernel 3.19.0-33. Non-surprisingly, we found that hyper-threaded 32 cores do not behave in the same manner as real 16 cores (numbers here for a single machine). We disregarded the hyper-threading ability in most of the
cases.

@@

\begin{table}
\caption{The benchmarks we use in this paper.}
\label{tab:benches}
\centering
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{0.5em}
\begin{tabular}{lccll}
\toprule
Name & Area & Type & Origin & Citation \\
\midrule
\rmtest & Mathematics & \ensuremath{\Varid{parMap}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\jacobitest & Mathematics & \ensuremath{\Varid{workpool}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\torustest & Mathematics & \ensuremath{\Varid{torus}} & Eden & \citet{Eden:SkeletonBookChapter02}\\
\sudokutest & Puzzle & \ensuremath{\Varid{parMap}} & \ensuremath{\Conid{Par}} Monad & \citet{par-monad} 
\tablefootnote{actual code from: http://community.haskell.org/~simonmar/par-tutorial.pdf and https://github.com/simonmar/parconc-examples}\\
\bottomrule
\end{tabular}
\end{table}

\rmtest is a probabilistic primality test that iterates multiple (32--256 here)
\enquote{subtests}. Should a subtest fail, the input is definitely not a
prime. If all $n$ subtest pass, the input is composite with the
probability of $1/4^{n}$.

@@

\begin{table}[]
\centering
\label{my-label}
\begin{subtable}[t]{.5\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{l R{1cm} R{1.3cm} R{1.3cm}}
		% \multicolumn{3}{c}{worst Sudoku Benchmarks} \\ \hline
		vs.             & cores	& params   & factor \\ \hline
		\csvreader[head to column names]{benchmarks/sudoku-sm/worstSudoku.csv}{}{\vs & \nCores & \params & \factor \\}
	\end{tabular}
	\endgroup
    \caption{Sudoku}\label{table1}
\end{subtable}
\caption{Worst Benchmark results}
\end{table}

@@

The PArrows version of \rmtest on a single node of the Glasgow grid
showed almost linear speedup (Fig.~\ref{fig:bench-rm-sm}). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65, the
efficiency was 85.3\%. However, if we increase the number of
requested cores to be 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup does not increase that well. It is merely 15.99
for 32 tasks with PArrows/Eden. It is worse for other backends.  As
for 64 tasks, we obtain the speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. Efficiency is 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
grid fails to execute two parallel Haskell processes with full-fledged
parallelism. For this reason, we did not regard hyper-threaded cores in
our speedup plots in Figs.~\ref{fig:bench-rm-sm}--\ref{fig:sudokuSMBenchmark}.

@@

\begin{figure}
%\centering
%\performanceSkelRMSM{11213}{64}\hfill%
{\speedupSkelRMSM{11213}{32}}\hfill%
{\speedupSkelRMSM{11213}{64}}
\caption{Relative speedup of \rmtest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware. Measurements were performed on a single node of the Glasgow
  grid; it has 16 real cores. Input was $2^{11213}-1$, we used 32 (left) or 64 (right)
  tasks. The
  closer to linear speedup the better.}
\label{fig:bench-rm-sm}
\end{figure}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSkelRMDist{44497}{256}{32}{272}\label{subfig:rm-dist-a}}%
%\hfill%
{\performanceSkelRMDistDiff{44497}{256}{32}{0.5}{272}\label{subfig:rm-dist-b}}
\caption{Parallel performance of the \rmtest on the Glasgow grid
  consisting of 256 cores. Input was $2^{44497}-1$, we used 256
  tasks. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:bench-rm-dist}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupJacobiDist{3}{3217}{32}{272}{4253}\label{subfig:jacobi-dist-a}}%
%\hfill%
{\performanceJacobiDistDiff{3}{3217}{32}{5}{272}\label{subfig:jacobi-dist-b}}
\caption{Parallel performance of the \jacobitest on the Glasgow grid
  consisting of 256 cores. Input was $2^{3217}-1$, we used 256
  tasks. The top plot shows relative speedup in a distributed memory setting compared to a simulated speedup for input $2^{4253}-1$. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:bench-jacobi-dist}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupTorusDist{4096}{32}{272}\label{subfig:speedupTorusDist}}%
%\hfill%
{\performanceTorusDistDiff{4096}{32}{0.5}{272}\label{subfig:performancetorusDistDiff}}
\caption{Parallel performance of \torustest on the Glasgow grid
  consisting of 256 cores. Input was a matrix size of $4096$. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:torusBenchmark}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{1000}\label{subfig:speedupSudokuSM}}%
%\hfill%
{\performanceSudokuSM{1000}\label{subfig:performanceSudokuSM}}
\caption{Absolute speedup of \sudokutest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the \ensuremath{\Varid{parMap}} version from the \ensuremath{\Conid{Par}} Monad examples. Measurements were performed on a single node of the Glasgow
  grid; it has 16 real cores and 32 threads. Input was a file of $1000$ Sudokus. The
  closer to linear speedup the better.}
\label{fig:sudokuSMBenchmark}
\end{figure}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{16000}\label{subfig:speedupSudokuSM16000}}%
%\hfill%
{\performanceSudokuSM{16000}\label{subfig:performanceSudokuSM16000}}
\caption{Absolute speedup of \sudokutest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the \ensuremath{\Varid{parMap}} version from the \ensuremath{\Conid{Par}} Monad examples. Measurements were performed on a single node of the Glasgow
  grid; it has 16 real cores and 32 threads. Input was a file of $16000$ Sudokus. The
  closer to linear speedup the better. The GpH version shows signs of choking with too many sparks being created.}
\label{fig:sudokuSMBenchmark16000}
\end{figure}

>>>>

\subsection{Hardware and Software used}

In the following we describe the benchmarks of our parallel DSL and algorithmic skeletons we conducted. They were run both in a shared and in a distributed
memory setting. All benchmarks were done on \enquote{the Glasgow GPG Beowulf cluster}, consisting of
16 machines with 2 Intel\SymbReg~Xeon\SymbReg~E5-2640 v2 and 64 GB of DDR3 RAM each. Each processor has 8 cores and 16 (hyper-threaded) threads with a base frequency of 2 GHz and a turbo frequency of 2.50 GHz. This results in a total of 256 cores and 512 threads for the whole cluster. The operating system was Ubuntu 14.04 LTS with Kernel 3.19.0-33. Non-surprisingly, we found that hyper-threaded 32 cores do not behave in the same manner as real 16 cores (numbers here for a single machine). We disregarded the hyper-threading ability in most of the cases.

Apart from Eden, all benchmarks and libaries were compiled with Stack's \footnote{see https://www.haskellstack.org/} lts-7.1 GHC compiler which is equivalent to a standard GHC 8.0.1 with the base package in version 4.9.0.0. Stack itself was used in version 1.3.2. For GpH in its Multicore variant we used the the parallel cabal package in version 3.2.1.0, while for the |Par| monad we used monad-par in version 0.3.4.8. For all Eden tests we used its GHC-Eden compiler in version 7.8.2 together with OpenMPI 1.6.5.

@@

\begin{table}
\centering
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{0.5em}
\begin{tabular}{lccll}
\toprule
Name & Area & Type & Origin & Citation \\
\midrule
\rmtest & Mathematics & \ensuremath{\Varid{parMap}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\jacobitest & Mathematics & \ensuremath{\Varid{workpool}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\torustest & Mathematics & \ensuremath{\Varid{torus}} & Eden & \citet{Eden:SkeletonBookChapter02}\\
\sudokutest & Puzzle & \ensuremath{\Varid{parMap}} & \ensuremath{\Conid{Par}} Monad & \citet{par-monad} 
\tablefootnote{actual code from: http://community.haskell.org/\~simonmar/par-tutorial.pdf and https://github.com/simonmar/parconc-examples}\\
\bottomrule
\end{tabular}
\label{tab:benches}
\caption{The benchmarks we use in this paper.}
\end{table}

\rmtest is a probabilistic primality test that iterates multiple (here: 32--256)
\enquote{subtests}. Should a subtest fail, the input is definitely not a
prime. If all $n$ subtest pass, the input is composite with the
probability of $1/4^{n}$.

@@

\input{bestAndWorstBenchmarks}

@@

The PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup (Fig.~\ref{fig:bench-rm-sm}). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65, the
efficiency was 85.3\%. However, if we increase the number of
requested cores to be 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup does not increase that well. It is merely 15.99
for 32 tasks with PArrows/Eden. It is worse for other backends.  As
for 64 tasks, we obtain the speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. Efficiency is 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
cluster fails to execute two parallel Haskell processes with full-fledged
parallelism. For this reason, we did not regard hyper-threaded cores in
our speedup plots in Figs.~\ref{fig:bench-rm-sm}--\ref{fig:sudokuSMBenchmark}.

@@

\begin{figure}
%\centering
%\performanceSkelRMSM{11213}{64}\hfill%
{\speedupSkelRMSM{11213}{32}}\hfill%
{\speedupSkelRMSM{11213}{64}}
\caption{Relative speedup of \rmtest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware. Measurements were performed on a single node of the Glasgow
  cluster; it has 16 real cores. Input was $2^{11213}-1$, we used 32 (left) or 64 (right)
  tasks. The
  closer to linear speedup the better.}
\label{fig:bench-rm-sm}
\end{figure}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSkelRMDist{44497}{256}{32}{272}\label{subfig:rm-dist-a}}%
%\hfill%
{\performanceSkelRMDistDiff{44497}{256}{32}{0.5}{272}\label{subfig:rm-dist-b}}
\caption{Parallel performance of the \rmtest on the Glasgow cluster
  consisting of 256 cores. Input was $2^{44497}-1$, we used 256
  tasks. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:bench-rm-dist}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupJacobiDist{3}{3217}{32}{272}{4253}\label{subfig:jacobi-dist-a}}%
%\hfill%
{\performanceJacobiDistDiff{3}{3217}{32}{5}{272}\label{subfig:jacobi-dist-b}}
\caption{Parallel performance of the \jacobitest on the Glasgow cluster
  consisting of 256 cores. Input was $2^{3217}-1$, we used 256
  tasks. The top plot shows relative speedup in a distributed memory setting compared to a simulated speedup for input $2^{4253}-1$. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:bench-jacobi-dist}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupTorusDist{4096}{32}{272}\label{subfig:speedupTorusDist}}%
%\hfill%
{\performanceTorusDistDiff{4096}{32}{0.5}{272}\label{subfig:performancetorusDistDiff}}
\caption{Parallel performance of \torustest on the Glasgow cluster
  consisting of 256 cores. Input was a matrix size of $4096$. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
higher the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:torusBenchmark}
\end{figure}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{1000}\label{subfig:speedupSudokuSM}}%
%\hfill%
{\performanceSudokuSM{1000}\label{subfig:performanceSudokuSM}}
\caption{Absolute speedup of \sudokutest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the \ensuremath{\Varid{parMap}} version from the \ensuremath{\Conid{Par}} Monad examples. Measurements were performed on a single node of the Glasgow
  cluster; it has 16 real cores and 32 threads. Input was a file of $1000$ Sudokus. The
  closer to linear speedup the better.}
\label{fig:sudokuSMBenchmark}
\end{figure}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{16000}\label{subfig:speedupSudokuSM16000}}%
%\hfill%
{\performanceSudokuSM{16000}\label{subfig:performanceSudokuSM16000}}
\caption{Absolute speedup of \sudokutest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the \ensuremath{\Varid{parMap}} version from the \ensuremath{\Conid{Par}} Monad examples. Measurements were performed on a single node of the Glasgow
  cluster; it has 16 real cores and 32 threads. Input was a file of $16000$ Sudokus. The
  closer to linear speedup the better. The GpH version shows signs of choking with too many sparks being created.}
\label{fig:sudokuSMBenchmark16000}
\end{figure}
